---
title: "Using the `carbon_view()` Shiny App"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the carbon_view() Shiny App}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

`carbon_view()` launches the built-in Shiny dashboard included with `tidycarbon`.
The app helps you inspect:

- System and R session details
- Machine configuration information detected through CodeCarbon
- Carbon and energy metrics from your `emissions.csv` file
- Carbon-equivalent context in a calculator panel

The dashboard is especially useful after you have tracked one or more runs and want a quick visual readout of impact.

## Before You Launch the App

You will get the best results if the following are ready:

1. `tidycarbon` is installed.
1. Python and the `codecarbon` module are available through `reticulate`.
1. At least one `emissions.csv` file exists in your project (the app looks for it recursively).

If no `emissions.csv` is found, the app will show:
`"emissions.csv not found anywhere in the project folder."`

## Create `emissions.csv` (if needed)

If you have not produced tracking output yet, run a quick tracked session:

```{r, eval = FALSE}
library(tidycarbon)

tracker <- carbon_init(project_name = "carbon_view_demo")
tracker_start(tracker)

# Example workload
Sys.sleep(1)
x <- sum(1:1000000)

tracker_stop(tracker)
```

After this, an `emissions.csv` file should be available for the dashboard.

## Launch the App

```{r, eval = FALSE}
library(tidycarbon)

# Opens in your browser by default
carbon_view()
```

You can also keep it in the R session viewer/browser behavior controlled manually:

```{r, eval = FALSE}
carbon_view(use_browser = FALSE)
```

## What You Will See

### 1) System Metrics

This tab contains:

- Core OS and R information (`Sys.info()` + `R.version`)
- Machine configuration fields gathered from CodeCarbon (CPU model, RAM size, region, etc.)
- Full `sessionInfo()` output

### 2) Carbon Analysis

This tab reads your emissions logs and shows:

- Total CO2 emissions (`emissions`)
- Total energy consumed (`energy_consumed`)
- Energy breakdown chart (`cpu_energy`, `gpu_energy`, `ram_energy`)
- A table view of raw metrics from the log

### 3) Carbon Equivalents

This tab displays your computed emissions value and embeds an external calculator view for context.

## Common Issues

If the app cannot find required columns in your `emissions.csv`, it will report messages such as:

- `"Column 'emissions' not found in emissions.csv."`
- `"Column 'energy_consumed' not found in emissions.csv."`
- `"Missing one or more columns: cpu_energy, gpu_energy, ram_energy."`

In practice, these are usually caused by:

1. Reading a non-CodeCarbon CSV with a matching name
1. A stale or manually edited emissions file
1. Mixed logs from different tools/formats

Use a fresh `tidycarbon` tracking run to regenerate a clean file when in doubt.

## Suggested Workflow

1. Track one or more scripts/functions with `tidycarbon`.
1. Open `carbon_view()` for exploratory diagnostics.
1. Compare emissions and energy totals after optimization changes.
1. Repeat and keep logs for transparent reporting.
